Economic Evaluation of Geographically Weighted Regression Analysis for Site-specific Nitrogen Management

### 2022-05-25 Writing/manuscript_gwr.rmd
bibliography updates

**Abstract**: Geographically weighted regression (GWR) has been presented as a prominent tool to estimate site-specific yield response functions to generate site-specific input rate recommendations using data generated by on-farm precision experimentation (OFPE). The key objectives of this paper are two-fold. First, it evaluates two sets of functional-form specifications made by two recent published studies of site-specific input management that apply GWR to data from OFPE. We show that one set of functional-form specification performs significantly better than the other under usual historical price scenarios. Second, we investigate the possibility of bias when GWR is used to estimate the value of site-specific vs. uniform-rate input management. We show that GWR suffers significant positive bias, which can lead to great over-estimation of the value of applying GWR methods. Our study is the first to systematically examine the statistical and economic properties of the use of GWR in OPFE settings. We reveal that in many circumstances using GWR is far less profitable than claimed by previous studies, which provides a necessary cautionary message to researchers who might otherwise “over-sell” to other researchers the value of using GWR methods in empirical work involving the estimation of site-specific input management.

@@

Second, we will examine the degree and direction of bias in the estimation of the value of GWR-based variable rate application (VRA) and question the extremely high estimates of that value reported by @trevisan2021spatial. In general, researchers take the following steps to estimate the value of VRA based using OFPE data [e.g., @swinton1998evaluating; @bullock2002adding; @anselin2004spatial; @liu2006site]: (1) estimate site-specific yield response functions based on a method chosen by the researcher; (2) use the estimated yield response functions to estimate the economically optimal site-specific input rates (EOIR) and economically optimal uniform input rates; (3) plug the estimated site-specific EOIRs and uniform EOIR into the estimated yield response functions to calculate estimated profits under both strategies; and (4) find the difference in the two strategies' estimated profits. However, this approach can result in significant bias in the estimation of the value of VRA because of the assumption made on step (3), which is that the estimated yield response functions are the true yield response functions. @trevisan2021spatial provide an example. At step (3), they used their estimated GWR as the true yield response functions. This assumption is generally unavoidable for researchers since they cannot observe any real field's true yield response function of any real field. Use of these procedures assumes accurate estimation of yield response functions, which when untrue can result in significant bias in the estimation of the value of VRA. This bias problem can be particularly severe when the method used tends to exaggerate the true heterogeneity in yield response functions. We suspect that the high value of VRA estimated by @trevisan2021spatial displays this bias.

@@

The field was 432m (72 cells) wide and 864m (144 cells) long, making its area 37.3 ha. Data from the 6-m-wide headlands, side-lands (there are no transition zones on the north and south ends of the field), and 12-m-wide "transition zones" were not included in the data analyzed. In total there were 1440 subplots available for experimental data analysis. The index $j \in {1, 2, \dots, 1440}$ identifies the field's subplots, and the index $k \in {1, \dots , 6}$ identifies the cells within a subplot. The couplet (j, k) is used to identify individual cells.

@@

An extensive literature examines the choice of functional form in yield response estimation [@cerrato1990comparison, @frank1990comparison, @bullock1994quadratic, @tembo2008crop]. The quadratic, linear-plateau, quadratic-plateau, and Mitscherlich‐Baule are commonly used functional forms. In this study, each cell (j, k)'s "true" yield response function is assumed to follow the quadratic-plateau functional form:

@@

The spatial distributions of cell-level parameters $\alpha_{j, k}$, $\beta_{j, k}$, $\gamma_{j,k}$, $\tau_{j, k}$, and the error term $\varepsilon_{j, k}$ were simulated by generating spatially correlated random fields using unconditional Gaussian geostatistical simulation based on the specified variograms [@dietrich1996fast, @wood1994simulation], and the simulation processes were conducted using the R package gstat [@graler16]. Specifically, the model is spherical, the nugget variance is zero, and the range is 600 meters. The nugget value of zero means that infinitesimally close observations have a zero covariance (i.e., they are identical). The range of 600m means that observations separated by more than 600m are not spatially correlated. A higher range value indicates increased spatial dependence, meaning less abrupt differences are observed between the production function parameters of neighboring cells. The mean and sill variance of each parameter were specified to result in a realistic mean yield of $11,416$ kg ha$^{-1}$, and such that the cells' true economically optimal N rate ranged from $113$ kg ha$^{-1}$ to $273$ kg ha$^{-1}$ among all the simulations, which most real-world N rate recommendations fall into (e.g., @sawyer2018). Note that the range of the optimal N rates of individual rounds of simulations are narrower. In defining the magnitude of the error term, we utilized the experimental data generated through eight on-farm randomized N field trials conducted by the DIFM project in 2018 by the Data-Intensive Farm Management Project (DIFM), [@bullock2019data]. Using the data from each of the eight actual trials, separate regressions of yield on N were run, and standard deviations of the residuals were calculated. The mean value of the eight standard deviations was $1370$ kg ha$^{-1}$, which was the value of sill variances assigned to the simulations' error terms.

@@

The GWR-R model is simply the basic GWR model (Fotheringham et al., 1996; Brunsdon et al., 1999; Fotheringham et al., 2002), where yield response was modeled as subplot-specific quadratic functions of the nitrogen rate, which for each subplot $j \in {1, 2, \dots, 1440}$ was specified as,

@@

The GWR-T model is also named as “mixed” GWR in the literature (Brunsdon et al., 1999) where some parameters are fixed globally while others vary geographically. In our model, the coefficient on $N^2$ was estimated as a global parameter while the coefficient on $N$ was estimated subplot-specifically. For each subplot $j \in {1, 2, \dots, 1440}$, the assumed yield response was,

@@

The Gaussian kernel function was used when estimating both GWR models, and the optimal bandwidth was selected through the corrected AIC approach (Fotheringham et al., 2002). The optimal bandwidth differed slightly across each simulated experiment, and the average bandwidth was around 18 meters.

@@

Figure \@ref(fig:pi-dif-dist) shows the distribution of $\hat{\Delta}_{gwr-m}$, profit difference between the GWR-based VRA and the SCAM-based URA. The distribution of the value of VRA from GWR-R is centered around 0, indicating that VRA based on GWR-R is on average no more profitable than SCAM-based URA. This result comes from GWR-R's highly inaccurate estimation of the subplot-level $N$ and $N^2$ coefficients, which led to unreliable subplot-specific nitrogen rate recommendations. Figure \@ref(fig:true-vs-estimated-coef-gwr-r) illustrates the accuracy of GWR-R in estimating the $N$ and $N^2$ coefficients in one simulation round. Estimates of the $N^2$ coefficients are positive for some subplots, indicating an increasing marginal product of the nitrogen rate, which contradicts a well-established agronomic property. The result is consistent with the bottom panel of Figure 12 in @rakshit2020 as well as the widely criticized poor coefficient estimation accuracy of GWR model in the literature (Griffith, 2008; Páez et al., 2011; Silva and Fotheringham, 2016; Harris, 2019). Figure \@ref(fig:true-vs-estimated-optn-gwr-r) plots the estimated cell-specific optimal N against the true optimal N for the same simulation round, and again shows that GWR-R can be highly unreliable, and consequently that basing variable rate management on GWR-R may be no better than taking a simple URA approach. Basically, not enough information is available locally to estimate reliably the coefficients on both $N$ and $N^2$ by subplot.

@@

Our simulation results show the basic GWR model (GWR-R) has little economic value for site-specific $R_xs$. The GWR-R based VRA is on average no more profitable than SCAM-based URA under historically common price scenarios. This finding reveals the economic consequences of the widely recognized statistical inaccuracy issue of GWR estimation (Griffith, 2008; Páez et al., 2011; Silva and Fotheringham, 2016; Harris, 2019), and in agreement with the previous critique that GWR should be used as an exploratory tool rather than an inferential tool (Wheeler and Tiefelsdorf, 2005; Páez et al., 2011). Precision agriculture practitioners should be cautious when using GWR for site-specific $R_xs$. If GWR-based approach must be employed to find site-specific input recommendations (e.g., due to lack of field and soil information), our simulation results show that GWR-T performs relatively better than GWR-R, though its profitability improvement against URA is only moderate.

@@

An obvious cause of the GWR models’ poor economic performances in this work is the quadratic functional form used in GWR that differs from the true quadratic-plateau yield response functions. One solution could be to upgrade to nonlinear GWR models. Lambert and Cho (2022) have recently developed a linear-plateau GWR model, which could be worthy of future examinations of its economic value. Another solution is to extend the parametric GWR approaches to a semi-parametric or non-parametric GWR where the yield response functions are estimated without assuming particular functional forms (like SCAM). However, such a method is likely to encounter a very similar problem to that faced by GWR-R: insufficient information available locally to estimate reliably the shape of yield response curve locally. It is also noted that the recently developed multiscale GWR model (Fotheringham et al., 2017; Oshan et al., 2019; Lu et al., 2019;)

[@yu2020inference; @comber2022route] allows the bandwidth to be derived separately for each covariate in the model, which would potentially improve the performances of GWR estimates and be incorporated into future research.

@@

A fruitful extension of GWR-based approach would be to employ more flexible yield reponse model specifications. This study has shown the assumed quadratic functional form to be the main culprit of the significant over-estimation of the value of GWR-based VRA. Applying GWR with a quadratic-plateau or linear-plateau specifications might substantially alleviate the reported bias problem.

>>>>

**Abstract**: Geographically weighted regression (GWR) has been presented as a prominent tool to estimate site-specific yield response functions to generate site-specific input rate recommendations using data generated by on-farm precision experimentation (OFPE). The key objectives of this paper are two-fold. First, it evaluates two sets of functional-form specifications made by two recent published studies of site-specific input management that apply GWR to data from OFPE. We show that one set of functional-form specification performs significantly better than the other under usual historical price scenarios _(Our trial design simulations show the two GWR models beat each other under different price scenarios.)_. Second, we investigate the possibility of bias _(Can we think of another term rather than “bias”? The term “bias” will tempt some reviewers (the not-so-smart ones) to think it is about the GWR coefficient estimation accuracy issue.)_ when GWR is used to estimate the value of site-specific vs. uniform-rate input management. We show that GWR suffers significant positive bias, which can lead to great over-estimation of the value of applying GWR methods. Our study is the first to systematically examine the statistical _(Maybe remove “statistical” to avoid any further attacks from statistician reviewers?)_ and economic properties of the use of GWR in OPFE settings. We reveal that in many circumstances using GWR is far less profitable than claimed by previous studies, which provides a necessary cautionary message to researchers who might otherwise “over-sell” to other researchers the value of using GWR methods in empirical work involving the estimation of site-specific input management.

@@

Second, we will examine the degree and direction of bias in the estimation of the value of GWR-based variable rate application (VRA) and question the extremely high estimates of that value reported by @trevisan2021spatial. In general, researchers take the following steps to estimate the value of VRA based using OFPE data [e.g., @anselin2004spatial; @swinton1998evaluating; @bullock2002adding;  @liu2006site]: (1) estimate site-specific yield response functions based on a method chosen by the researcher; (2) use the estimated yield response functions to estimate the economically optimal site-specific input rates (EOIR) and economically optimal uniform input rates; (3) plug the estimated site-specific EOIRs and uniform EOIR into the estimated yield response functions to calculate estimated profits under both strategies; and (4) find the difference in the two strategies' estimated profits. However, this approach can result in significant bias in the estimation of the value of VRA because of the assumption made on step (3), which is that the estimated yield response functions are the true yield response functions. @trevisan2021spatial provide an example. At step (3), they used their estimated GWR as the true yield response functions. This assumption is generally unavoidable for researchers since they cannot observe any real field's true yield response function of any real field. Use of these procedures assumes accurate estimation of yield response functions, which when untrue can result in significant bias in the estimation of the value of VRA. This bias problem can be particularly severe when the method used tends to exaggerate the true heterogeneity in yield response functions. We suspect that the high value of VRA estimated by @trevisan2021spatial displays this bias.

@@

The field was 432m (72 cells) wide and 864m (144 cells) long, making its area 37.3 ha. Data from the 6-m-wide headlands, side-lands (there are no transition zones on the north and south ends of the field), and 12-m-wide "transition zones" were not included in the data analyzed. In total there were 1440 subplots available for experimental data analysis. The index $j \in {1, 2, \dots, 1440}$ identifies the field's subplots, and the index $k \in {1, \dots , 6}$ identifies the cells within a subplot. The couplet $(j, k)$ is used to identify individual cells.

@@

An extensive literature examines the choice of functional form in yield response estimation [@cerrato1990comparison, @frank1990comparison, @bullock1994quadratic, @tembo2008crop]. The quadratic, linear-plateau, quadratic-plateau, and Mitscherlich‐Baule are commonly used functional forms. _(How about we just focus on quadratic-plateau, which is well documented in the agronomy literature. So no reviewers will keep asking to do "robustness checks" of other functional forms.)_ In this study, each cell $(j, k)$'s "true" yield response function is assumed to follow the quadratic-plateau functional form: _(I tried adding equation number by using the "\tag{}". It works in .rmd, but won't be knitted into .docx file.)_

@@

The spatial distributions of cell-level parameters $\alpha_{j, k}$, $\beta_{j, k}$, $\gamma_{j,k}$, $\tau_{j, k}$, and the error term $\varepsilon_{j, k}$ were simulated by generating spatially correlated random fields using unconditional Gaussian geostatistical simulation based on the specified variograms [@dietrich1996fast, @wood1994simulation], and the simulation processes were conducted using the R package gstat [@graler16]. Specifically, the model is spherical, the nugget variance is zero, and the range is 600 meters. The nugget value of zero means that infinitesimally close observations have a zero covariance (i.e., they are identical). The range of 600m means that observations separated by more than 600m are not spatially correlated. A higher range value indicates increased spatial dependence, meaning less abrupt differences are observed between the production function parameters of neighboring cells. The mean and sill variance of each parameter were specified to result in a realistic mean yield of $11,416$ kg ha$^{-1}$, and such that the cells' true economically optimal N rate ranged from $113$ kg ha$^{-1}$ to $273$ kg ha$^{-1}$ among all the simulations, which most real-world N rate recommendations fall into [e.g., @sawyer2018]. Note that the range of the optimal N rates of individual rounds of simulations are narrower. In defining the magnitude of the error term, we utilized the experimental data generated through eight on-farm randomized N field trials conducted by the DIFM project in 2018 by the Data-Intensive Farm Management Project (DIFM), [@bullock2019data]. Using the data from each of the eight actual trials, separate regressions of yield on N were run, and standard deviations of the residuals were calculated. The mean value of the eight standard deviations was $1370$ kg ha$^{-1}$, which was the value of sill variances assigned to the simulations' error terms.

@@

The GWR-R model is simply the basic GWR model [@fotheringham1996geography; @brunsdon1999some; @fotheringham2003geographically], where yield response was modeled as subplot-specific quadratic functions of the nitrogen rate, which for each subplot $j \in {1, 2, \dots, 1440}$ was specified as,

@@

The GWR-T model is also named as “mixed” GWR in the literature [@brunsdon1999some] where some parameters are fixed globally while others vary geographically. In our model, the coefficient on $N^2$ was estimated as a global parameter while the coefficient on $N$ was estimated subplot-specifically. For each subplot $j \in {1, 2, \dots, 1440}$, the assumed yield response was,

@@

The Gaussian kernel function was used when estimating both GWR models, and the optimal bandwidth was selected through the corrected AIC approach [@fotheringham2003geographically]. The optimal bandwidth differed slightly across each simulated experiment, and the average bandwidth was around 18 meters.

@@

Figure \@ref(fig:pi-dif-dist) shows the distribution of $\hat{\Delta}_{gwr-m}$, profit difference between the GWR-based VRA and the SCAM-based URA. The distribution of the value of VRA from GWR-R is centered around 0, indicating that VRA based on GWR-R is on average no more profitable than SCAM-based URA. This result comes from GWR-R's highly inaccurate estimation of the subplot-level $N$ and $N^2$ coefficients, which led to unreliable subplot-specific nitrogen rate recommendations. Figure \@ref(fig:true-vs-estimated-coef-gwr-r) illustrates the accuracy of GWR-R in estimating the $N$ and $N^2$ coefficients in one simulation round. Estimates of the $N^2$ coefficients are positive for some subplots, indicating an increasing marginal product of the nitrogen rate, which contradicts a well-established agronomic property. The result is consistent with the bottom panel of Figure 12 in @rakshit2020 as well as the widely criticized poor coefficient estimation accuracy of GWR model in the literature [@griffith2008spatial; @paez2011simulation; @da2016multiple; @harris2019simulation]. Figure \@ref(fig:true-vs-estimated-optn-gwr-r) plots the estimated cell-specific optimal N against the true optimal N for the same simulation round, and again shows that GWR-R can be highly unreliable, and consequently that basing variable rate management on GWR-R may be no better than taking a simple URA approach. Basically, not enough information is available locally to estimate reliably the coefficients on both $N$ and $N^2$ by subplot.

@@

Our simulation results show the basic GWR model (GWR-R) has little economic value for site-specific $R_xs$. The GWR-R based VRA is on average no more profitable than SCAM-based URA under historically common price scenarios. This finding reveals the economic consequences of the widely recognized statistical inaccuracy issue of GWR estimation [@griffith2008spatial; @paez2011simulation; @da2016multiple; @harris2019simulation], and in agreement with the previous critique that GWR should be used as an exploratory tool rather than an inferential tool [@wheeler2005multicollinearity; @paez2011simulation]. Precision agriculture practitioners should be cautious when using GWR for site-specific $R_xs$. If GWR-based approach must be employed to find site-specific input recommendations (e.g., due to lack of field and soil information), our simulation results show that GWR-T performs relatively better than GWR-R, though its profitability improvement against URA is only moderate.

@@

An obvious cause of the GWR models’ poor economic performances in this work is the quadratic functional form used in GWR that differs from the true quadratic-plateau yield response functions. One solution could be to upgrade to nonlinear GWR models. @lambert2022geographically have recently developed a linear-plateau GWR model, which could be worthy of future examinations of its economic value. Another solution is to extend the parametric GWR approaches to a semi-parametric or non-parametric GWR where the yield response functions are estimated without assuming particular functional forms (like SCAM). However, such a method is likely to encounter a very similar problem to that faced by GWR-R: insufficient information available locally to estimate reliably the shape of yield response curve locally. It is also noted that the recently developed multiscale GWR model [@fotheringham2017multiscale; @yu2020inference; @comber2022route] allows the bandwidth to be derived separately for each covariate in the model, which would potentially improve the performances of GWR estimates and be incorporated into future research.

@@

A fruitful extension of GWR-based approach would be to employ more flexible yield reponse model specifications. This study has shown the assumed quadratic functional form to be the main culprit of the significant over-estimation of the value of GWR-based VRA. Applying GWR with a quadratic-plateau or linear-plateau _(Since Lambert has already developed the linear-plateau model, how about we delete it here?)_ specifications might substantially alleviate the reported bias problem.
